Reviewer Report

### 2018-12-31 docs/reviewer_comments.Rmd
Made last changes and re-submitted

- There are some rather loose statements in the ms, starting with the Abstract (see also Rev 1) where it is implied that a 'well-managed' fishery should result in no effect in a reserve. This is false. Most 'well-managed' fisheries aim to bring B0 to around 50% of unfished biomass, so reserves should on average contain double the density of fished areas. There is one unsubstantiated statement (line 81) of good enforcement, but could it not be just as likely that there is fishing activity in these reserves? Poaching is an equally valid explanation as any examine din the Discussion.

*That is correct, a reserve would contain higher density, biomass, and richness. We have corrected the abstract to reflect what we intended to say, which is that a well-managed fishery would not see increased catches by an additional intervention. We do not include explicit references in the abstract, but the discussion touches on this topic and cites Hilborn, Micheli and De Leo's paper that evaluates when MRs result in increased catches.*

*As for the unsubstantiated statement of good enforcement, it is unlikely that these reserves experience poaching. The reserves are implemented, managed, and monitored by the community themselves. We have expanded on this when we introduce the case studies.*

@@

*We have expanded on the methods, especially by adding to the description of case studies (TURFs, reserve implementation process) and data collection methods to clarify what the sampling design is. We have also expanded on the regression model. It is a multiple linear regression because we incorporate more than one explanatory variable. The lambda that we report does not make reference to the lambda parameter in a Poisson distribution, but rather to the coefficient presented in Equations 1 and 2. We do not use a log-linear model that assumes a discrete poison distribution because our indicators are density (i.e. org / m^-2^) and biomass (Kg / m^-2^), which are continuous variables.*

@@

*The observation that Moland's paper might have confounded estimates of reserve effectiveness is correct. Proximity of their reserve and control sites might imply a spillover effect from the reserves to the controls may occur, which would result in more conservative estimates of reserve effectiveness. While there is a long history of time series analyses in ecology, we cite Moland's paper because, to our knowledge, it is the first paper that estimated reserve effectiveness using a difference-in-differences approach. We have included references to older implementations of BACI in ecology, as well as the seminal econometric paper that introduces the Difference-in-Differences methodology.*

@@

*A particular characteristic of Fish Refuges is that they are implemented with a pre-determined expiration date. Reserves are only re-implemented if the fishers decide to do so. We view this willingness to renew the reserves as a continued support. We have expanded in the description of the case studies and discussion to explain the reserve implementation process and why we think that these reserves continue to be supported. Furthermore, the work by Ayer et al., 2018 explicitly looked at the perceptions that these communities have about these reserves, and their findings support our statements. We have modified the last sentence of the manuscript to present a solid conclusion about the recommendations for the reserves we evaluate, but without speculating about fisher's perception of marine conservation.*

@@

*The points raised above were amongst the most useful comments, and the fact that these were highlighted by both reviewers suggest that there was room for improvement. We have significantly expanded the introduction by including more background on TURFs, reserves, and TURF reserves. We have re-written some paragraphs to highlight the fact that TURF-reserve effectiveness has been proposed from theoretical and modelling work, but that no empirical evidence exists for reserves*

@@

*We have included more information on the wider context, touching on implementation processes and TURF history.*

@@

*We have toned down the generalizations and made distinctions between our claims of when to generalize. Specifically, we meant to say that one may generalize our findings to other TURF-managed fisheries where community-based reserves are implemented. Further down in the discussion, we meant to say that no generalizations should be made to other Fish Refuges not implemented in TURF-managed fisheries. We believe that our current description and expanded introduction help make this more clear.*

@@

*We have included more information on data collection as supplementary materials, where we show the number of transects for each site (reserve - control) and type of survey (invertebrate - fish). The process of site selection is presented in the methods, where we expanded the definition of our control sites (L142)*

@@

*We have explicitly mentioned the objective of evaluating the SES in our methods section (L134), and further explained variable selection in L203.*

@@

*We have expanded on the methods section and further explained the Difference-in-Differences method (L178). We have also included a reference to the main paper (Card and Krueger, 1994) where the DiD methodology was first introduced.*

@@

*Our goal is to report the effect of the reserve on the indicator, and using effect sizes allows us to reduce the ambiguity that temporal and spatial trends may induce. However, we recognize the importance of observing "raw" data, and have included time series of all indicators in the supplementary materials.*

@@

*The interpretation is correct, we intend to provide enough information to show that the characteristics of the system are conductive of effective resource management. The reason why similar broad descriptions are presented across sites is because these don't vary across them. For example, all sites must align with the same official fisheries regulation (Lobster Fishery Official Mexican Norm; NOM-006-pesc-1993). Our end goal is to present the information as a contextual summary. The use of the SES framework simply allows us to be more objective by being transparent about the variables we report across communities and using a common language. We believe this is now made clear in sentence starting in L134.*

@@

*We believe that part of the disconnect had to do with the way in which the previous text was presented, where we failed to distinguish between the evaluated TURF-reserves, other non-evaluated TURF-reserves and other reserves not implemented within TURFs. We have also made it clear that some of the claims (which include citations) come from conclusions of other works performed on these same sites. For example, the buffering to environmental variation comes from Micheli et al., 2012, who evaluated the TURF-reserves in Isla Natividad.*

@@

*We have included a paragraph on some of these limitations in the discussion (L271)*

@@

*Figures and table have been modified or added to the manuscript.*

@@

*We have expanded on the methods and expect these to be now better documented.*

@@

*Our data and code are deposited in a public GitHub Repository. We have included a sentence in the methods sections mentioning this and a link to said repository (L201).*

@@

*The social evaluation is, indeed, secondary to this study. While we believe it is equally important, we assess reserve effectiveness based on the stated objectives, which were to increase lobster densities and landings. We incorporate the SES approach to provide more background about the system under which the reserves are placed, and to discard the usually-mentioned causes of lack of effectiveness which "blame" fishers (namely poaching and lack of enforcement). Instead, our approach shows that the most likely cause of lack of effectiveness is poor design, that ignore basic design principles such as size of reserves.*

@@

*We believe that the re-structuring of the introduction and case study descriptions have now addressed this.*

@@

*No fishing activities occur within the Marine Reserves, and we have now made it clear that these are no-take marine reserves.*

@@

*Yes, these fisheries have been historically managed with TURFs. We have included this in the case study description (sentence starting in L120).*

@@

*We were using the term "triple bottom line evaluation" in the same way as Halpern 2013 did. We realize that this is not clear, and have now explicitly stated that we "provide a holistic evaluation of the effectiveness of community-based in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop" (L65).*

@@

*The sizes were included in Table 2 (RS3 - Size of resource system: reserve area), but are now also mentioned in the results (L248). We do not directly assess the effect of reserve size because all reserves are smaller than the suggested design principles (Green et al., 2017). If we had data from more sites and across a wider range of reserve sizes (i.e. smaller and greater than minimum requirements) it would be interesting to show how time to recovery or effect sizes relate to reserve size. One would expect to find a positive correlation.*

@@

*Yes, all control sites are placed within the TURFs, as it is now explicitly mentioned in the case study section (list starting in L141).*

@@

*Mean ex-vessel prices for lobster are now presented in Table S3*

@@

*Yes, our Difference-in-Differences analyses controls for community-specific changes.*

@@

*Differences in revenue may be explained by both, a simple increase in prices due to market interactions or an increase in landings (plausibly because of the reserve). However, we do not observe an increase in revenues nor in landings, and ex-vessel prices remain relatively constant.*

@@

*We did not directly evaluate the community- or species-level effects on the indicators. Since we fit one model for each combination of indicator-community, we cannot make inference about the community- or species-level effects. This type of analyses would be more suited for a community ecology approach, or when seeking differential response among species (e.g. Babcock et al, 2010). Furthermore, we believe that this approach should be used in a post hoc fashion for reserves that have actually had an effect on the indicators, seeking to understand which species and communities may be driving this.*

@@

*This assumption is correct, we only list the number of fishing co-op members. However, all fishers that fish within the TURF belong to the cooperative, and there are no "free fishers". This point is now mentioned in the case study section (L90).*

@@

*We have included the linear distance to the closes urban centers in Table 2. Isla Natividad lies 545 Km south from Tijuana, while Maria Elena and Punta Herrero are more than 200 km south from Cancun.*

@@

*Yes, each community engages in biological monitoring (the data for this study comes from this) and oceanographic monitoring.*

@@

*We believe that splitting the narrative column into three would be beneficial if each community had very different values / narratives. However, these three communities have strikingly similar values. Splitting it into three columns would result in redundant text being repeated across columns. However, we do understand the point that it is difficult to recognize which values correspond to each community. To address this, we have modified the table caption to explain that the values apply to all communities, and when needed a specific community-level narrative is included (for example, isolation of IN vs PH and ME).*

@@

*This is a fair point, which was raised previously. This would be an interesting approach. Initially, we had "operationalized" the SES framework and included a scoring system. However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them. We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores. The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management. We have included a discussion paragraph recognizing this (L330).*

@@

*This big fluctuations in biomass are explained by an opportunistic increase in sizes. In this particular case, a few larger-than-usual organisms were observed, but the total number of fish remained similar to other years. As a result, the few larger fish observed these years yield a higher biomass. However these differences are not significantly different from the pre-observed trend.*

@@

*After the first few years of reserve implementation, one would expect landings to decrease. This would be explained by the relative reduction of total fishable area. The "dip" described in Ovando et al., (2016) would then recover as densities inside the reserve recover and spillover occurs. The fact that we don't see a reduction in landings over the first few years suggests that reserves are not big enough to reduce landings over the first few years.*

@@

*We have removed this lose statement and included an explicit conclusions paragraph (L335) where we state that "We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside".*

@@

*The previous wording of this statement was confusing, and seemed like we were referencing the reserves evaluated in this work. Clearly this is not true and our data would not support such claim. As with previous statements, we believe this is now clear after re-structuring the discussion.*

@@

place-holder

*This would be an interesting approach. Initially, we had "operationalized" the SES framework and included a scoring system. However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them. We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores. The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management. We have included a discussion paragraph recognizing this (L330).*

>>>>

- There are some rather loose statements in the ms, starting with the Abstract (see also Rev 1) where it is implied that a 'well-managed' fishery should result in no effect in a reserve. This is false. Most 'well-managed' fisheries aim to bring B0 to around 50% of unfished biomass, so reserves should on average contain double the density of fished areas. There is one unsubstantiated statement (line 81) of good enforcement, but could it not be just as likely that there is fishing activity in these reserves? Poaching is an equally valid explanation as any examined in the Discussion.

*That is correct, a reserve would contain higher density, biomass, and richness. We have corrected the abstract to reflect what we intended to say, which is that a well-managed fishery would not see increased aggregate catches by an additional intervention. We do not include explicit references in the abstract, but the discussion touches on this topic and cites Hilborn, Micheli and De Leo's paper that evaluates when MRs result in increased catches.*

*As for the unsubstantiated statement of good enforcement, it is unlikely that these reserves experience poaching. The reserves are implemented, managed, and monitored by the communities themselves. We have expanded on this when we introduce the case studies, and our SES approch suggests high levels of self-enforcement to protect the TURF and reserves.*

@@

*We have expanded on the methods, especially by adding to the description of case studies (TURFs, reserve implementation process) and data collection methods to clarify what the sampling design is. We have also expanded on the regression models. It is a multiple linear regression because we incorporate more than one explanatory variable (time, zone, and the interaction between these). The lambda that we report does not make reference to the lambda parameter in a Poisson distribution, but rather to the coefficient presented in Equations 1 and 2. We do not use a log-linear model that assumes a discrete poison distribution because our indicators are density (i.e. org / m^-2^) and biomass (Kg / m^-2^), which are continuous variables.*

*Additionally, we realized that our equations were incorrectly formulated. They included species- and community-level fixed effects. However, our code did not incorporate these terms, as our data does not allow us to include them. The equations and text have been modified to reflect our actual analysis.*

@@

*The observation that Moland's paper might have confounded estimates of reserve effectiveness is correct. Proximity of their reserve and control sites might imply a spillover effect from the reserves to the controls, which would result in downard-biased estimates of reserve effectiveness. While there is a long history of time series analyses in ecology, we cite Moland's paper because of the methods, not the results; to our knowledge, it is the first paper that estimated reserve effectiveness using a difference-in-differences approach. Furthermore, we have included references to older implementations of BACI in ecology, as well as the seminal econometric paper that introduces the difference-in-differences methodology.*

@@

*A particular characteristic of Fish Refuges is that they are implemented with a pre-determined expiration date. Reserves are only re-implemented if the fishers decide to do so. We view this willingness to re-implement the reserves as an explicit manifestation of support. We have expanded in the description of the case studies and discussion to explain the reserve implementation process and why we think that these reserves continue to be supported, which may help support our point. Furthermore, the work by Ayer et al., 2018 explicitly looked at the perceptions that these communities have about these reserves, and their findings support our statements. We have modified the text to present a solid conclusion about the recommendations for the reserves we evaluate, but without speculating about fisher's perception of marine conservation.*

@@

*The points raised above were amongst the most useful comments, and the fact that these were highlighted by both reviewers suggest that there was room for improvement on our side. We have significantly expanded the introduction by including more background on and distinctions between TURFs, reserves, and TURF reserves. We have re-written parts of the introduction to highlight the fact that TURF-reserve effectiveness has been proposed from theoretical and modelling work, but that no empirical evidence exists for reserves*

@@

*We have included more information on the wider context, touching on implementation processes and TURF history.We have also expanded on the way in which the evaluated reserves are implemented, which may further help the reader understand TURF systems and TURF-reserves in Mexico.*

@@

*We have toned down the generalizations and made distinctions between our claims of when to generalize. Specifically, we meant to say that one may generalize our findings to other TURF-managed fisheries where community-based reserves are implemented. Further down in the discussion, we meant to say that no generalizations should be made to other Fish Refuges not implemented in TURF-managed fisheries. We believe that our current description and expanded introduction help make this more clear. In recognition of the caution needed when generalizing, we have also re-structured the title to make reference to Mexican small-scale fisheries.*

@@

*We have included more information on data collection as supplementary materials, where we show the number of transects for each site (reserve - control) and type of survey (invertebrate - fish). The process of site selection is presented in the methods, where we expanded on the definition of our control sites and their location (L135).*

@@

*We have explicitly mentioned the objective of evaluating the SES in our methods section (L122), and further explained variable selection in L197.*

@@

*We have expanded on the methods section and further explained the Difference-in-Differences approach (L170). We have also included a reference to the main paper (Card and Krueger, 1994) where the DiD methodology was first introduced in econometric literature.*

@@

*Our goal is to report the effect of the reserves on each indicator, and using effect sizes allows us to reduce the ambiguity that temporal and spatial trends may induce. However, we recognize the importance of observing "raw" data, and have included time series of all indicators in the supplementary materials.*

@@

*The interpretation is correct, we intend to provide enough information to show that the characteristics of the system are conductive of effective resource management. The reason why similar broad descriptions are presented across sites is because there is little variation between them. For example, all sites must align with the same official fisheries regulation (Lobster Fishery Official Mexican Norm; NOM-006-pesc-1993). Our end goal is to present the information as a contextual summary. The use of the SES framework simply allows us to be more objective by being transparent about the variables we report across communities and using a common language. We believe this is now made clear in sentence starting in L129.*

@@

*We believe that part of the disconnect had to do with the way in which the previous text was presented, where we failed to distinguish between the evaluated TURF-reserves and other hypothetical TURF-reserves. We have now made it clear that some of the claims come from conclusions of other works performed on these same sites; we have included relevant references. For example, the buffering to environmental variation comes from Micheli et al., 2012, who evaluated the TURF-reserves in Isla Natividad.*

*The point raised about the possible use of resources for other interventions is interesting. This issue has been partially addressed by other literature, such as Smith et al (2010) on what economists call "Opportunity costs". We have included a mention of this in the discussion (L327).*

@@

*We have included a paragraph on some of these limitations in the discussion (L264)*

@@

*The requested figures and tables have been added to the Supplementary Materials.*

@@

*We have expanded on the methods.*

@@

*Our data and code are deposited in a public GitHub Repository. We have included a sentence in the methods sections mentioning this and a link to said repository (L195). The same information is available within the supplementary materials and in the "data and code archiving" section of the submission.*

@@

*The social evaluation is, indeed, secondary to this study. While we believe it is equally important, we assess reserve effectiveness based on the stated objectives, which were to increase lobster densities and landings. We incorporate the SES approach to provide more background and context about the system under which the reserves are placed, and to systematically address the usually-mentioned causes of lack of effectiveness which "blame" fishers (namely poaching and lack of enforcement). Instead, our approach shows that the most likely cause of lack of effectiveness is poor design (i.e. reserves are too small).*

*We agree that an operationalization of the SES that allows us to match scores to effectiveness would be an interesting approach. Initially, we had "operationalized" the SES framework and included a scoring system. However, since all sites operate under very similar settings, scores were practically the same and obtained little contrast between them. We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores. Furthermore, they all seem to show a common cause for lack of effectivness. The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management. We have included this in our discussion (L330).*

@@

*We believe that the re-structuring of the introduction and case study descriptions have provided additional context needed (Paragraphs in L39 and further in L77 and L85)..*

@@

*No fishing activities occur within the Marine Reserves, and we have now made it clear that these are no-take marine reserves (L34 and L135).*

@@

*Yes, these fisheries have been historically managed with TURFs. We have included this in the case study description (sentence starting in L115).*

@@

*We were using the term "triple bottom line evaluation" in the same way as Halpern 2013 did. We realize that this was not a clear term. Instead, we have now explicitly stated that we "provide a holistic evaluation of the effectiveness of community-based in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop" (L60).*

@@

*The sizes were included in Table 2 (RS3 - Size of resource system: reserve area), but are now also mentioned in the results (L243). We do not directly assess the effect of reserve size because all reserves are smaller than the suggested design principles (Green et al., 2017). If we had data from more sites and across a wider range of reserve sizes (i.e. smaller and greater than minimum requirements) it would be interesting to show how time to recovery and effect sizes relate to reserve size. One would expect to find a positive correlation between these, but that plateaus as one reaches optimum size.*

@@

*Yes, all control sites are placed within the TURFs, as it is now explicitly mentioned in the case study section (list starting in L135).*

@@

*These are now presented in Table S3*

@@

*Yes, our Difference-in-Differences analyses controls for these changes.*

@@

*Differences in revenue may be explained by both, an increase in prices due to market interactions or an increase in landings (plausibly because of the reserve or increased effort). However, we do not observe an increase in revenues or in landings, and ex-vessel prices remain relatively constant.*

@@

*We did not directly evaluate the community- or species-level effects on the indicators. Since we fit one model for each combination of indicator-community, we cannot make inference about the community- or species-level effects. This type of analysis would be more suited for a community ecology approach, or when seeking differential response among species (e.g. Babcock et al, 2010). Furthermore, we believe that this approach should be used in a post hoc fashion for reserves that have actually had an effect on the indicators, seeking to understand which species and communities may be driving this. We have included part of this explanation on model fitting when the analyses section was re-structured.*

@@

*This assumption is correct, we only list the number of fishing co-op members. However, all fishers that fish within the TURF belong to the cooperative, and there are no unaffiliated fishers. This point is now mentioned in the case study section (L85).*

@@

*We have included the linear distance to the closest urban centers in Table 2. Isla Natividad lies 545 Km south from Tijuana, while Maria Elena and Punta Herrero are more than 200 km south from Cancun. Both cities have more than 1 Million people, international airports, and receive large amounts of tourism.*

@@

*Yes, each community engages in biological monitoring  and oceanographic monitoring programmes. In fact, the data for this study comes from such monitoring efforts.*

@@

*We agree that splitting the narrative column into three would be beneficial if each community had very different values / narratives. However, these three communities have strikingly similar values. Splitting it into three columns would result in redundant text being repeated across columns. However, we do understand the point that it is difficult to recognize which values correspond to each community. To address this, we have modified the table caption to explain that the values apply to all communities, and when needed a specific community-level narrative is included (for example, isolation of IN vs PH and ME).*

@@

*This is a fair point, which was raised previously. This would be an interesting approach if we were opreationalizing the framework. As before, we state: Initially, we had "operationalized" the SES framework and included a scoring system. However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them. We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores. The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management. We have included a discussion paragraph recognizing this (L330).*

@@

*This big fluctuations in biomass are explained by an increase in sizes. In this particular case, a few larger-than-usual organisms were observed, but the total number of fish remained similar to other years. As a result, the few larger fish observed these years yield a higher biomass. However these differences are not significantly different from the pre-observed trend or from other years.*

@@

*After the first few years of reserve implementation, one would expect landings to decrease. This would be explained by the relative reduction of total fishable area. The "dip" described in Ovando et al., (2016) would then recover as densities inside the reserve increase and spillover occurs. The fact that we don't see a reduction in landings over the first few years suggests that reserves are not big enough to reduce landings over the first few years.*

@@

*We have removed this lose statement and included an explicit conclusions paragraph (L322) where we state that "We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside".*

@@

*Our previous wording of this statement made it confusing, and seemed like we were referencing the reserves evaluated in this work. Clearly this is not true and our data would not support such claim. As with previous statements, we believe this is now clear after re-structuring the discussion.*

@@
